\sysfull

### 2018-12-27 paper/TOSN/background.tex
TOSN coala version

\begin{table}
    \centering
    \footnotesize
    \begin{tabular}{|c|c|}
        \hline
        Model & Data Copied to/from NVRAM \\
        \hline\hline
        Mementos~\cite{mementos}    & Registers + Stack     \\
        DINO~\cite{dino}    & Registers + Stack + WAR NV variables \\%used in task\\
        Chain~\cite{chain}  & PC + NV variables used in task\\
        Alpaca~\cite{alpaca}    & PC + WAR NV variables used in task\\
        Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017} & Registers (requires NV main memory) \\
        Region Formation~\cite{baghsorkhi_cgo_2018} & Registers + Updated variables in task \\
        \hline
    \end{tabular}
    \caption{Non-volatile memory access for data consistency; {PC}: program counter, {WAR variables}: variables involved in Write-after-Read (WAR) dependencies, {NV}: non-volatile.}
    \label{table:chechpoint_comparison}
\end{table}

@@

\begin{figure}
	\begin{subfigure}{0.49\columnwidth}
		\includegraphics[width=\columnwidth]{figures/war-example.pdf}
		\caption{WAR-affected code}
		\label{fig:war-example}
	\end{subfigure}
	\begin{subfigure}{0.49\columnwidth}
		\includegraphics[width=\columnwidth]{figures/war-execution.pdf}
		\caption{WAR-affected execution}
		\label{fig:war-execution}
	\end{subfigure}
	\caption{WAR dependency example. \texttt{NV} marks non-volatile variables, \texttt{checkpoint()} is a checkpoint of volatile state.	\label{fig:w-a-r}}
\end{figure}

>>>>

% Table
\begin{table}%
\caption{Simulation Configuration}
\label{tab:one}
\begin{minipage}{\columnwidth}
\begin{center}
\begin{tabular}{ll}
  \toprule
        Model & Data Copied to/from NVRAM \\
        \hline
        Mementos~\cite{mementos}    & Registers + Stack     \\
        DINO~\cite{dino}    & Registers + Stack + WAR NV variables \\%used in task\\
        Chain~\cite{chain}  & PC + NV variables used in task\\
        Alpaca~\cite{alpaca}    & PC + WAR NV variables used in task\\
        Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017} & Registers (requires NV main memory) \\
        Region Formation~\cite{baghsorkhi_cgo_2018} & Registers + Updated variables in task \\
  \bottomrule
\end{tabular}
\end{center}
\end{minipage}
  \label{table:chechpoint_comparison}
\end{table}%

%\begin{table}
%    \centering
%    \begin{tabular}{|c|c|}
%        \hline
%        Model & Data Copied to/from NVRAM \\
%        \hline
%        Mementos~\cite{mementos}    & Registers + Stack     \\
%        DINO~\cite{dino}    & Registers + Stack + WAR NV variables \\%used in task\\
%        Chain~\cite{chain}  & PC + NV variables used in task\\
%        Alpaca~\cite{alpaca}    & PC + WAR NV variables used in task\\
%        Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017} & Registers (requires NV main memory) \\
%        Region Formation~\cite{baghsorkhi_cgo_2018} & Registers + Updated variables in task \\
%        \hline
%    \end{tabular}
%    \caption{Non-volatile memory access for data consistency; {PC}: program counter, {WAR variables}: variables involved in Write-after-Read (WAR) dependencies, {NV}: non-volatile.}
%    \label{table:chechpoint_comparison}
%\end{table}

@@

\begin{figure}
	\begin{subfigure}{0.4\columnwidth}
		\centering
		\includegraphics[width=\columnwidth]{figures/war-example.pdf}
		\caption{WAR-affected code}
		\label{fig:war-example}
	\end{subfigure}
	\begin{subfigure}{0.4\columnwidth}
		\centering
		\includegraphics[width=\columnwidth]{figures/war-execution.pdf}
		\caption{WAR-affected execution}
		\label{fig:war-execution}
	\end{subfigure}
	\caption{WAR dependency example. \texttt{NV} marks non-volatile variables, \texttt{checkpoint()} is a checkpoint of volatile state.	\label{fig:w-a-r}}
\end{figure}

### 2018-12-27 paper/TOSN/compiler.tex
TOSN coala version

To sub-divide these tasks, the compiler prototype does depth-first traversals
of all paths in the CFG, ignoring back-edges. Traversals start from the
beginning of the main function, or from an existing task boundary. The
prototype accumulates the cost of basic blocks during the traversal until
reaching another task boundary or an exit node in the CFG. At the end of a
traversal, if the accumulated cost exceeds to estimated device energy capacity the
compiler prototype must split the traversed path with a task boundary. Each
edge on the path is a candidate location for a boundary.

\sys uses a heuristic to compute a score for each candidate and places a
boundary at the candidate with the highest score. The heuristic tries to avoid
placing boundaries on loops, tries to minimize the number of live variables 
protected across the new boundary, and tries to maximize the number of operations
along the path. A boundary candidate's score is computed as
%
\begin{equation}
\text{score} = {\left(w_{0} N_{\text{IR}} - w_{1} N_{\text{PV}}\right)}{N_{\text{loop}}^{-1}},\nonumber
\end{equation}
%
where $N_{\text{IR}}$ is the length of the path being traversed up to the
candidate, $N_{\text{PV}}$ is the number of protected variables that will be
newly introduced by placing boundary at the candidate, and $N_{\text{loop}}$ is
the product of the loop bounds of all loops (to handle nested loops) containing the
boundary candidate. \sys normalizes $N_{\text{IR}}$ and $N_{\text{PV}}$ to the
values among boundary candidates. The weights, $w_{0}, w_{1}$ are between zero
and one, and are hand-tuned parameters that define the relative importance of
$N_{\text{IR}}$, $N_{\text{PV}}$, and $N_{\text{loop}}$. We empirically determined these weight values by measuring the runtime for different weight values across
applications. We found that varying weights varies runtime by only around 5\%
in most cases and in our experimental evaluation, we used $w_{0} = w_{1} =
0.5$.

>>>>

To sub-divide these tasks, the compiler prototype does depth-first traversals of all paths in the CFG, ignoring back-edges. Traversals start from the beginning of the main function, or from an existing task boundary. Our prototype accumulates the cost of basic blocks during the traversal until reaching another task boundary or an exit node in the CFG. At the end of a traversal, if the accumulated cost exceeds to estimated device energy capacity the compiler prototype must split the traversed path with a task boundary. Each edge on the path is a candidate location for a boundary.

\textbf{Boundary Placement Heuristic.} \sys uses a heuristic to compute a score for each candidate and places a boundary at the candidate with the highest score. The heuristic tries to avoid placing boundaries on loops, tries to minimize the number of live variables protected across the new boundary, and tries to maximize the number of operations along the path. A boundary candidate's score is computed as
%
\begin{equation}
\text{score} = {\left(w_{0} N_{\text{IR}} - w_{1} N_{\text{PV}}\right)}{N_{\text{loop}}^{-1}},\nonumber
\end{equation}
%
where $N_{\text{IR}}$ is the length of the path being traversed up to the
candidate, $N_{\text{PV}}$ is the number of protected variables that will be
newly introduced by placing boundary at the candidate, and $N_{\text{loop}}$ is
the product of the loop bounds of all loops (to handle nested loops) containing the
boundary candidate. \sys normalizes $N_{\text{IR}}$ and $N_{\text{PV}}$ to the
values among boundary candidates. The weights, $w_{0}, w_{1}$ are between zero
and one, and are hand-tuned parameters that define the relative importance of
$N_{\text{IR}}$, $N_{\text{PV}}$, and $N_{\text{loop}}$. We empirically determined these weight values by measuring the runtime for different weight values across
applications. We found that varying weights varies runtime by only around 5\%
in most cases and in our experimental evaluation, we used $w_{0} = w_{1} =
0.5$.

### 2018-12-27 paper/TOSN/implementation.tex
TOSN coala version

We implemented \sys's programming and execution model as a runtime library and API that a programmer can use to make a plain C program intermittency-safe.
%
\subsection{Application Programming Interface}
\label{sec:coala_api}
%
Coala's API adds only a few syntactic constructs to a C-based language, summarized
in Table~\ref{table:coala_api}.
%
\begin{table}
\footnotesize
\centering
\begin{tabular}{| r | p{0.7\columnwidth} |}
	\hline
	{Method} & {Arguments} \\
	\hline\hline
	\texttt{INIT}($t$) & $t \in \mathbf{T}$: scheduled task on first boot \\
	\hline
	\texttt{RUN}() & --- \\
	\hline
	\texttt{TASK}($t$, $w_t$) & $t \in \mathbf{T}$: task name, $w_t$: weight of task $t$ \\
	\hline
	\texttt{NEXT\_TASK}($t$) & $t \in \mathbf{T}$ : task to be run next \\
	\hline
	\texttt{PV}($p$, $v$ [, $s$]) & $p$: type, $v \in \mathbf{V}$: name, $s$: array size \\
	\hline
	$u$ := \texttt{RP}($v$) & $v \in \mathbf{V}$: protected variable to read, $u$: dest.\ operand \\
	\hline	
	\texttt{WP}($v$) := $u$ &  $v \in \mathbf{V}$: protected variable to write, $u$: source operand \\
	\hline
	\texttt{SM}($p$, $m$ [, $s$]) & $p$: type, $m$: name, $s$: array size \\
	\hline
	\texttt{DISABLE\_PC}() & --- \\
	\hline
	\texttt{ENABLE\_PC}() & --- \\
	\hline
\end{tabular}
\caption{API summary; $\mathbf{T}$: set of all tasks, $\mathbf{V}$: set of all protected variables, $[, s]$: optional argument.}
\label{table:coala_api}
\end{table}

\paragraph{New Tasks.} The \texttt{TASK} annotation on a function
declaration statically allocates a non-volatile constant variable holding a
task's weight and declares that the function is a task.

\paragraph{Task Transitions.} \texttt{NEXT\_TASK} marks the task to be executed at the 
end of the current one task and it can be invoked along
any control path to dynamically determine the next task.

\paragraph{Protected Variables.} The \texttt{PV} annotation on a
variable statically allocates a protected non-volatile variable. The
variable must then be accessed with the \texttt{RP} and \texttt{WP} API methods at
run-time to ensure correct operation.

\paragraph{Initialization.} The behavior of the API method
\texttt{INIT} is very similar to \texttt{NEXT\_TASK}, with the
addition of performing preliminary kernel initializations, including hardware setup.

\paragraph{Execution.} The programmer passes control to \sys's task scheduler 
by calling \texttt{RUN} after device initialization. 
%
First the scheduler updates the coalescing target using the coalescing
strategy.  Second, the scheduler resumes a commit, if one is in progress. Third,
the scheduler clears the list of dirty shadow pages. The scheduler then 
sets the program counter to the next task to run, which \sys tracks in  
non-volatile memory.  
%
Before executing the task, \sys checks whether there is a partially committed
task to resume, which requires \sys to restore volatile state, including the program
counter.  
%
If there is no in-progress, partially committed task, \sys starts executing and
coalescing tasks.

>>>>

We implemented \sys's programming and execution model as a runtime library and API that a programmer can use to make a plain C program intermittency-safe.
%
\subsection{Application Programming Interface}
\label{sec:coala_api}
%
Coala's API adds only a few syntactic constructs to a C-based language, summarized
in Table~\ref{table:coala_api}.
%
\begin{table}%
\caption{Simulation Configuration}
\label{tab:one}
\begin{minipage}{\columnwidth}
\begin{center}
\begin{tabular}{ll}
  \toprule
	{Method} & {Arguments} \\
	\hline
	\texttt{INIT}($t$) & $t \in \mathbf{T}$: scheduled task on first boot \\
	\texttt{RUN}() & --- \\
	\texttt{TASK}($t$, $w_t$) & $t \in \mathbf{T}$: task name, $w_t$: weight of task $t$ \\
	\texttt{NEXT\_TASK}($t$) & $t \in \mathbf{T}$ : task to be run next \\
	\texttt{PV}($p$, $v$ [, $s$]) & $p$: type, $v \in \mathbf{V}$: name, $s$: array size \\
	$u$ := \texttt{RP}($v$) & $v \in \mathbf{V}$: protected variable to read, $u$: dest.\ operand \\
	\texttt{WP}($v$) := $u$ &  $v \in \mathbf{V}$: protected variable to write, $u$: source operand \\
	\texttt{SM}($p$, $m$ [, $s$]) & $p$: type, $m$: name, $s$: array size \\
	\texttt{DISABLE\_PC}() & --- \\
	\texttt{ENABLE\_PC}() & --- \\
  \bottomrule
\end{tabular}
\caption{API summary; $\mathbf{T}$: set of all tasks, $\mathbf{V}$: set of all protected variables, $[, s]$: optional argument.}
\label{table:coala_api}
\end{center}
\end{minipage}
\end{table}%
%%
%\begin{table}
%\footnotesize
%\centering
%\begin{tabular}{| r | p{0.7\columnwidth} |}
%	\hline
%	{Method} & {Arguments} \\
%	\hline\hline
%	\texttt{INIT}($t$) & $t \in \mathbf{T}$: scheduled task on first boot \\
%	\hline
%	\texttt{RUN}() & --- \\
%	\hline
%	\texttt{TASK}($t$, $w_t$) & $t \in \mathbf{T}$: task name, $w_t$: weight of task $t$ \\
%	\hline
%	\texttt{NEXT\_TASK}($t$) & $t \in \mathbf{T}$ : task to be run next \\
%	\hline
%	\texttt{PV}($p$, $v$ [, $s$]) & $p$: type, $v \in \mathbf{V}$: name, $s$: array size \\
%	\hline
%	$u$ := \texttt{RP}($v$) & $v \in \mathbf{V}$: protected variable to read, $u$: dest.\ operand \\
%	\hline	
%	\texttt{WP}($v$) := $u$ &  $v \in \mathbf{V}$: protected variable to write, $u$: source operand \\
%	\hline
%	\texttt{SM}($p$, $m$ [, $s$]) & $p$: type, $m$: name, $s$: array size \\
%	\hline
%	\texttt{DISABLE\_PC}() & --- \\
%	\hline
%	\texttt{ENABLE\_PC}() & --- \\
%	\hline
%\end{tabular}
%\caption{API summary; $\mathbf{T}$: set of all tasks, $\mathbf{V}$: set of all protected variables, $[, s]$: optional argument.}
%\label{table:coala_api}
%\end{table}

\noindent\textit{New Tasks.} The \texttt{TASK} annotation on a function
declaration statically allocates a non-volatile constant variable holding a
task's weight and declares that the function is a task.

\noindent\textit{Task Transitions.} \texttt{NEXT\_TASK} marks the task to be executed at the 
end of the current one task and it can be invoked along
any control path to dynamically determine the next task.

\noindent\textit{Protected Variables.} The \texttt{PV} annotation on a
variable statically allocates a protected non-volatile variable. The
variable must then be accessed with the \texttt{RP} and \texttt{WP} API methods at
run-time to ensure correct operation.

\noindent\textit{Initialization.} The behavior of the API method
\texttt{INIT} is very similar to \texttt{NEXT\_TASK}, with the
addition of performing preliminary kernel initializations, including hardware setup.

\noindent\textit{Execution.} The programmer passes control to \sys's task scheduler 
by calling \texttt{RUN} after device initialization.

First the scheduler updates the coalescing target using the coalescing
strategy.  Second, the scheduler resumes a commit, if one is in progress. Third,
the scheduler clears the list of dirty shadow pages. The scheduler then 
sets the program counter to the next task to run, which \sys tracks in  
non-volatile memory.  
%
Before executing the task, \sys checks whether there is a partially committed
task to resume, which requires \sys to restore volatile state, including the program
counter.  
%
If there is no in-progress, partially committed task, \sys starts executing and
coalescing tasks.

### 2018-12-27 paper/TOSN/relatedwork.tex
TOSN coala version

\textbf{Intermittently-powered Devices.} There is a large body of research on energy-harvesting and ambient-powered embedded devices summarized by~\cite{prasad_comst_2014,sample_procieee_2013,huang:commag:2015,visser_procieee_2013,kamalinejad_commag_2015,ku_cst_2016}. For instance, a new wave of embedded systems powered by radio waves is emerging~\cite{patel_pervasive_2017,rf_powered_computing_gollakota_2014,wisp5,moo,zhao_rfid_2015,holleman_biocas_2008,thomas_jbcs_2012,naderiparizi_rfid_2015,rodriguez_tbcs_2015,liu_sigcomm_2013,kicksat,nadeau_naturebio_2017}. The emergence of such hardware platforms
has led to the development of instrumentation, debugging, and prototyping tools for such systems
~\cite{hester_sensys_2014,hester_sensys_2015,edb,capybara,stork,wisent,flicker}.

\textbf{Checkpointing-based Systems.} Early work on energy-harvesting runtimes, such as Dewdrop~\cite{dewdrop}, assumed simple computations that complete on a predictable burst of energy.
%
%Eon~\cite{sorber_sensys_2007} has scheduled such computation based on priority.
%
Support for computation that spans power failures was first achieved with statically-placed conditional checkpoints in Mementos~\cite{mementos}. 
%Mementos left the possibility of non-volatile state becoming inconsistent with volatile state saved in the checkpoint, rendering in-place writes to data structures in non-volatile memory unsafe~\cite{mspcdino}. 
DINO~\cite{dino} addressed the consistency problem by selectively versioning non-volatile state within the checkpoints. Ratchet~\cite{ratchet} ensured consistency by placing a checkpoint at the beginning of each idempotent region in the code. Ratchet has a similar technique to task splitting with the core difference: the checkpoint is of a fixed size, while \sys privatize variant number of memory pages.
%
%Ratchet has a similar technique to task splitting with some essential differences: (i) It does not need to handle non-volatile memory accesses, since it assumes a unified non-volatile memory, whereas \sys has to save the volatile pages in non-volatile memory; (ii) The checkpoint is of a fixed size, while \sys privatize variant number of memory pages. 
%
Clank~\cite{hicks_isca_2017} ensured consistency with custom hardware that dynamically tracks read-after-write dependencies in memory accesses and checkpoints on demand. The consistency problem was also approached with a combination of undo and redo logging in software~\cite{baghsorkhi_cgo_2018}. Just-in-time checkpointing, such as Quickrecall~\cite{quickrecall} and Hibernus++~\cite{hibernusplusplus}, eschews inconsistency by saving all volatile state immediately before power failure and halting the execution. Unlike \sys, such systems rely on introspection hardware to monitor supply voltage and on accurate worst-case bounds on checkpoint cost.

In all of the above systems, with the exception of DINO~\cite{dino}, checkpoints are dynamic, i.e. the programmer does not have explicit control over the point at which the code may be resumed after power failure. Dynamic checkpointing systems make it difficult for the programmer to respect application-level atomicity constraints, such as correlating sensor readings. Checkpointing systems that copy most volatile state scale poorly as the size of volatile memory increases.  QuickRecall, Clank and Ratchet reduce the copying overhead by allocating the stack in non-volatile memory---requires more time and energy to access than volatile memory (cf. Section~\ref{sec:cost_task-based}) and not a viable option for off-chip non-volatile memory. In addition, QuickRecall and Clank require custom hardware.

\textbf{Task-based Systems.} An alternative to checkpointing are recent systems based on static tasks, such as Chain~\cite{chain}, Alpaca~\cite{alpaca} and Mayfly~\cite{hester_sensys_2017}. Using static tasks, they eliminate the need to checkpoint volatile state. Using channel-based memory models~\cite{chain,hester_sensys_2017} or automatic privatization and redo-logging~\cite{alpaca} they avoid checkpointing overheads. Moreover, task-based models facilitate respecting application-level atomicity constraints. \sys also relies on
statically defined tasks to avoid checkpointing volatile state. However, unlike prior systems, \sys coalesces its statically-defined tasks at runtime into more efficient dynamic tasks that adapt to changing energy conditions. \sys's mechanism for ensuring memory consistency also differs from the channel-based~\cite{chain} and privatization-based~\cite{alpaca} mechanisms in prior systems. \sys keeps memory consistent through memory virtualization optimized for bulk accesses to task-shared data with high locality.

\textbf{Task Decomposition.} In contrast to \sys's construction of coalesced tasks at runtime, prior work has proposed to optimize task size at compile time. CleanCut~\cite{cleancut_2018} program analysis statically estimated energy consumption and split the program into tasks until all tasks consumed less energy than the device can store. An alternative program analysis~\cite{baghsorkhi_cgo_2018} generated different versions of a program with different task sizes and empirically selected the best among them~\cite{baghsorkhi_cgo_2018}. HarvOS~\cite{mottola2017harvos} takes a hybrid approach that uses a program analysis to place a minimal number of conditional checkpoints that test the energy level at runtime before copying state, like Mementos~\cite{mementos}. Unlike \sys, such compiler-based approaches, face the challenge of statically predicting energy consumption of arbitrary input-dependent code with peripheral access, which is an problem without a general solution. Furthermore, a static decomposition approach prevents portability across devices with different storage capacitors. In contrast, \sys avoids forcing any assumptions at compile time and adapts to energy storage capacity and incoming energy conditions at runtime.

\textbf{Memory Virtualization.} Prior work on embedded systems has studied a variety of memory virtualization strategies relating to \sys.
TinyOS~\cite{levis2005tinyos} and nesC~\cite{nesc} support dynamic memory management. Later work extended the memory manager to support memory virtualization backed by flash memory~\cite{sensornetvm} and to ensure memory and type safety~\cite{tinyosmemorysafety}. SOS~\cite{sos}, Contiki~\cite{contiki}, and t-kernel~\cite{tkernel} also developed memory management abstractions that virtualize memory size and provide safe and indirect access. Mat\'e~\cite{mate} developed full virtual machines support for sensor nodes, virtualizing not just memory resources, but other states and peripherals. The goal of \sys is to provide consistent, intermittent execution, leveraging the benefits of efficient bulk copying. In contrast, prior efforts focused more on programmability and runtime reliability properties provided by virtual memory.

A related domain is unbounded, page-based transactional memory and deterministic parallel runtime systems~\cite{pagebasedtm,grace}.  These works have a different mechanism and purpose than \sys---ensuring that data are consistent and deterministically updated during concurrent executions. Their similarity with \sys is in managing state to ensure consistency at the granularity of pages to amortize checking and tracking costs. \sys's paging implementation, which keeps a shadow page for each page to use during commit, is similar to the shadow paging scheme use for transactional commit~\cite{pagebasedtm}.

>>>>

\paragraph{Intermittently-powered Devices.}
There is a large body of research on energy-harvesting and ambient-powered embedded devices summarized by~\cite{prasad_comst_2014,sample_procieee_2013,huang:commag:2015,visser_procieee_2013,kamalinejad_commag_2015,ku_cst_2016}. For instance, a new wave of embedded systems powered by radio waves is emerging~\cite{patel_pervasive_2017,rf_powered_computing_gollakota_2014,wisp5,moo,zhao_rfid_2015,holleman_biocas_2008,thomas_jbcs_2012,naderiparizi_rfid_2015,rodriguez_tbcs_2015,liu_sigcomm_2013,kicksat,nadeau_naturebio_2017}. The emergence of such hardware platforms
has led to the development of instrumentation, debugging, and prototyping tools for such systems
~\cite{hester_sensys_2014,hester_sensys_2015,edb,capybara,stork,wisent,flicker}.

\paragraph{Checkpointing-based Systems.}
Early work on energy-harvesting runtimes, such as Dewdrop~\cite{dewdrop}, assumed simple computations that complete on a predictable burst of energy.
%
%Eon~\cite{sorber_sensys_2007} has scheduled such computation based on priority.
%
Support for computation that spans power failures was first achieved with statically-placed conditional checkpoints in Mementos~\cite{mementos}. 
%Mementos left the possibility of non-volatile state becoming inconsistent with volatile state saved in the checkpoint, rendering in-place writes to data structures in non-volatile memory unsafe~\cite{mspcdino}. 
DINO~\cite{dino} addressed the consistency problem by selectively versioning non-volatile state within the checkpoints. Ratchet~\cite{ratchet} ensured consistency by placing a checkpoint at the beginning of each idempotent region in the code. Ratchet has a similar technique to task splitting with a core difference: the checkpoint is of a fixed size, while \sys privatizes a varying number of memory pages.
%
%Ratchet has a similar technique to task splitting with some essential differences: (i) It does not need to handle non-volatile memory accesses, since it assumes a unified non-volatile memory, whereas \sys has to save the volatile pages in non-volatile memory; (ii) The checkpoint is of a fixed size, while \sys privatize variant number of memory pages. 
%
Clank~\cite{hicks_isca_2017} ensured consistency with custom hardware that dynamically tracks WAR dependencies in memory accesses and checkpoints on demand. The consistency problem was also approached with a combination of undo- and redo-logging in software~\cite{baghsorkhi_cgo_2018}. Just-in-time checkpointing, such as Quickrecall~\cite{quickrecall} and Hibernus++~\cite{hibernusplusplus}, eschews inconsistency by saving all volatile state immediately before power failure and halting the execution. Unlike \sys, such systems rely on introspection hardware to monitor supply voltage and on accurate worst-case bounds on checkpoint cost.

In all of the above systems, with the exception of DINO~\cite{dino}, checkpoints are dynamic, i.e., the programmer does not have explicit control over the point at which the code may be resumed after power failure. Dynamic checkpointing systems make it difficult for the programmer to respect application-level atomicity constraints, such as correlating sensor readings. Checkpointing systems that copy most volatile state scale poorly as the size of volatile memory increases.  QuickRecall, Clank and Ratchet reduce the copying overhead by allocating the stack in non-volatile memory---requires more time and energy to access than volatile memory (cf. Section~\ref{sec:cost_task-based}) and not a viable option for off-chip non-volatile memory. In addition, QuickRecall and Clank require custom hardware.

\paragraph{Task-based Systems.}
Alternatives to checkpointing are recent systems based on static tasks, such as Chain~\cite{chain}, Alpaca~\cite{alpaca} and Mayfly~\cite{hester_sensys_2017}. Using static tasks, they eliminate the need to checkpoint volatile state. Using channel-based memory models~\cite{chain,hester_sensys_2017} or automatic privatization and redo-logging~\cite{alpaca} they avoid checkpointing overheads. Moreover, task-based models facilitate respecting application-level atomicity constraints. \sys also relies on
statically defined tasks to avoid checkpointing volatile state. However, unlike prior systems, \sys coalesces its statically-defined tasks at runtime into more efficient dynamic tasks that adapt to changing energy conditions. \sys's mechanism for ensuring memory consistency also differs from the channel-based~\cite{chain} and privatization-based~\cite{alpaca} mechanisms in prior systems. \sys keeps memory consistent through memory virtualization optimized for bulk accesses to task-shared data with high locality.

\paragraph{Task Decomposition.}
In contrast to \sys's construction of coalesced tasks at run-time, prior work has proposed to optimize task size at compile-time. CleanCut~\cite{cleancut_2018} program analysis statically estimates energy consumption and splits the program into tasks until all tasks consume less energy than the device can store. An alternative program analysis generates different versions of a program with different task sizes and empirically selects the best among them~\cite{baghsorkhi_cgo_2018}. HarvOS~\cite{mottola2017harvos} takes a hybrid approach that uses a program analysis to place a minimal number of conditional checkpoints that test the energy level at run-time before copying state, like Mementos~\cite{mementos}. Unlike \sys, such compiler-based approaches face the challenge of statically predicting energy consumption of arbitrary input-dependent code with peripheral access, which is a problem without a general solution. Furthermore, a static decomposition approach prevents portability across devices with different storage capacitors. In contrast, \sys avoids forcing any assumptions at compile time and adapts to energy storage capacity and incoming energy conditions at run-time.

\paragraph{Memory Virtualization.}
Prior work on embedded systems has studied a variety of memory virtualization strategies relating to \sys.
TinyOS~\cite{levis2005tinyos} and nesC~\cite{nesc} support dynamic memory management. Later work extended the memory manager to support memory virtualization backed by flash memory~\cite{sensornetvm} and to ensure memory and type safety~\cite{tinyosmemorysafety}. SOS~\cite{sos}, Contiki~\cite{contiki}, and t-kernel~\cite{tkernel} also developed memory management abstractions that virtualize memory size and provide safe and indirect access. Mat\'e~\cite{mate} developed full virtual machine support for sensor nodes, virtualizing not just memory resources, but other states and peripherals. The goal of \sys is to provide consistent, intermittent execution, leveraging the benefits of efficient bulk copying. In contrast, prior efforts focused more on programmability and run-time reliability properties provided by virtual memory.

A related domain is unbounded, page-based transactional memory and deterministic parallel runtime systems~\cite{pagebasedtm,grace}.  These works have a different mechanism and purpose than \sys---ensuring that data are consistent and deterministically updated during concurrent executions. Their similarity with \sys is in managing state to ensure consistency at the granularity of pages to amortize checking and tracking costs. \sys's paging implementation, which keeps a shadow page for each page to use during commit, is similar to the shadow paging scheme used for transactional commit~\cite{pagebasedtm}.
